### 2021-04-17 com/c1.tex
WIP 17Apr2021

>>>>

\item Given that $U$ is open, $C$ is compact, $C \subset U$ so that, $U^c$ is 
closed $C \cap U^c = \varnothing$. Therefore, (from the previous exercise),
there exits $d > 0$ such that for all $x \in C, y \in U^c$, $\norm{x - y} \ge
d$.

Define the collection $\{B_{d/2}(x) | x \in U\}$. It is an open cover of $U$
and $C \subset U$. Since $C$ is compact, there is finite subcover, say $
B_{d/2}(x_1) \cup \cdots \cup B_{d/2}(x_n)$ that covers $C$.

Let 
\[
D = \overline{\cup_{i=1}^n B_{d/2}(x_i)}
\]
then clearly $C \subset D \subset U$. Further, if $X$ is an open set then 
$X$ lies in the interior of $\overline{X}$.
\end{enumerate}

\section{Functions and continuity}\label{c1s3}
\begin{defn}\label{c1s3d1}
Consider a function $f: \sor^n \rightarrow \sor^m$. Then
\[
\lim_{x \rightarrow a}f(x) = b
\]
if for any $\epsilon > 0$, there exists $\delta > 0$ such that for all $0 < 
\norm{x - a} \le \delta$, $\norm{f(x) -b} \le \epsilon$.
\end{defn}
\begin{rem}
$f$ need not be defined at $x = a$ for its limit to be defined at $x = a$.
\end{rem}

\begin{defn}\label{c1s3d2}
Consider a function $f: \sor^n \rightarrow \sor^m$. $f$ is said to be continuous
at $x = a$ if for any $\epsilon > 0$, there exists $\delta > 0$ such that 
$\norm{x - a} \le \delta \Rightarrow \norm{f(x) - f(a)} \le epsilon$.
\end{defn}
\begin{rem}
$f$ must be defined at $x = a$ for it to be continuous at $x = a$.
\end{rem}

\begin{thm}\label{c1s3t1}
If $A \subset \sor^n$ and $f: A \rightarrow \sor^m$. Then $f$ is continuous
on $A$ if and only if for every open set $U \subset \sor^m$, there is an open 
set $V \subset \sor^n$ sich that $f^{-1}(U) = V \cap A$.
\end{thm}
\begin{proof}
Assume that $f$ is continuous in $A$. Let $U$ be an open set in $\sor^m$.
Let $a \in f^{-1}(U)$. By continuity of $f$ in A, for any $\epsilon > 0$, there
exists $\delta > 0$ such that $x \in B_\delta(a) \Rightarrow f(x) \in B_\epsilon
(f(a))$. Now the open ball $B_\delta(a)$ need not lie entirely in $A$. Therefore, we say that for any $B_\epsilon(f(a))$ there exists a set $A \cap B_\delta(a)$
such that $x \in A \cap B_\delta(a) \Rightarrow f(x) \in B_\epsilon(f(a))$. This
is true for all points $a \in A$.

To prove the converse, let $U = B_\epsilon(y)$, where $y \in \sor^m$. Then
there exists an open set $V \subset \sor^n$ such that $f^{-1}(U) = V \cap A$.
Choose $\delta$ such that $B_\delta(f^{-1}(y)) \subset V \cap A$. This proves
the continuity of $f$.
\end{proof}

\begin{rem}
The greatest significance of this theorem is that one can gets an equivalent
definition of continuity that does not require limits.
\end{rem}

Continuous functions have several useful properties.
\begin{thm}\label{c1s3t2}
If $f: A \rightarrow \sor^m$ is continuous where $A \subset \sor^n$ and $A$
is compact then $f(A)$ is also compact.
\end{thm}
\begin{proof}
Since $A$ is compact, it has a finite open cover. Since $f$ is continuous, it
maps an open set in $A$ to an open set in $f(A)$. Thus, $f(A)$ can be covered
with a finite number of open sets.
\end{proof}

\begin{rem}
What about the converse? Functions can be many-to-one. Continuity guarantees
that the pre-image is open but it cannot guarantee that it is finite. $f(x) =
\tanh(x)$, $x \in \sor$ is continuous but $\atanh([-1, 1])$ is not compact.
\end{rem}

\begin{defn}\label{c1s3d3}
Let $f: A \rightarrow \sor$ be bounded. If
\begin{eqnarray*}
M(a, f, \epsilon) &=& \sup\{f(x) : x \in B_\epsilon(a) \cap A\} \\
m(a, f, \epsilon) &=& \inf\{f(x) : x \in B_\epsilon(a) \cap A\} 
\end{eqnarray*}
then the oscillation of $f$ at $a$ is
\[
o(f, a) = \lim_{\epsilon \rightarrow 0} (M(a, f, \epsilon) - m(a, f, \epsilon)).
\]
\end{defn}

\begin{lem}\label{c1s3l1}
$o(f, a)$ always exists.
\end{lem}
\begin{proof}
Since $f$ is bounded, the infimum and supremum exist.
\end{proof}

\begin{thm}\label{c1s3t3}
A bounded function $f$ is continuous at $a$ if and only if $o(f, a) = 0$.
\end{thm}
\begin{proof}
If $f$ is continuous at $a$ then $f(x) \in B_{\epsilon}(f(a))$ or $M(x, f, a)
= f(a) + \epsilon$ and $m(x, f, a) = f(a) - \epsilon$ from which it follows
that $o(f, a) = 0$.

Conversely, if $o(f, a) = 0$ then as $\delta \rightarrow 0$, $M(a,f,\delta)
- m(a, f, \delta) < \epsilon \Rightarrow f(x) \in (f(a) - \epsilon/2, f(a) + 
\epsilon/2)$.
\end{proof}

\subsection{Problems}
\begin{enumerate}
\item If 
\[
\lim_{x \rightarrow a} f^i(x) = b^i
\]
for all $i = 1, \ldots, m$ then it is clear that $f(x) = (f^1(x), \ldots,
f^m(x)) \rightarrow (b^1, \ldots, b^m) = b$ as $x \rightarrow a$.

To prove the converse, note that 
\[
\norm{f(x) - b}^2 = \sum_{i=1}^m (f^i(x) - b^i)^2 < \epsilon^2 \Rightarrow
\abs{f^i(x) - b^i} < \epsilon.
\]

\item If $f^i$ is continuous at $a$ then $\abs{f^i(x) - f^i(a)} < \epsilon/
\sqrt{m}$ and
\[
\norm{f(x) - f(a)}^2 = \sum_{i=1}^m \abs{f^i(x) - f^i(a)}^2 < \epsilon^2
\Rightarrow \norm{f(x) - f(a)} < \epsilon.
\]

Conversely, 
\[
\norm{f(x) - f(a)}^2 < \epsilon^2 \Rightarrow \abs{f^i(x) - f^i(a)}
^2 < \epsilon^2 \Rightarrow \abs{f^i(x) - f^i(a)} < \epsilon.
\]

\item Let $T: \sor^n \rightarrow \sor^m$ be a linear transformation. Consider
$T(x + h) - T(x) = T(h) \le M\norm{h}$, by problem 10 of section \ref{c1s1}.
Therefore, if $\norm{h} < \epsilon/M$ then $\norm{T(x + h) - T(x)} < \epsilon$.

\item The set $A$ is the region in the first quadrant between the $x$-axis and
the parabola $y = x^2$.
\begin{enumerate}
\item A straightline that does not cut through $A$ surely has an interval
aroung the origin in $R^2 - A$ because the entire line lies in $R^2 - A$. A 
line that cuts $A$ still has an interval around the origin in $R^2 - A$
because for small $x$, $y = x^2 < mx$ and therefore the line is entirely in $R^2
- A$.

\item $f$ is the indicator function for $A$. For $h \in \sor^2$, define 
$g_h: \sor \rightarrow \sor = f(th)$. Thus, $g_h(t) = 1$ if $th \in A$ else it
is zero. To show that $g_h(t)$ is continuous at $t = 0$, consider an interval
$(-\delta, \delta)$. Any line through the origin has an interval around the 
origin entirely in $R^2 - A$ on which $g_h(t)$ takes the value $0$. Therefore,
$g_h$ as a function of $t$ is continuous at the origin.

However $f(h)$ is not continuous because it takes a value $1$ if we approach
$0$ from the positive $x$ axis and it takes a value $0$ if we approach from
any other side.
\end{enumerate}

\item The function $f(x) = \norm{x - a}$ is continuous.

\item $A$ is not closed. Therefore, there is at least one limit point of $A$,
say $x_0$, that is not part of $A$. The function $f = \norm{x - x_0}^{-1}$ is
an example of a continuous but unbounded function on $A$.

\item Since $A$ is compact and $f: A \rightarrow \sor$ is continuous, $f(A)$
is compact. Since $f(A) \subset \sor$, its compactness means that $f(A)$ is 
also closed and bounded. Therefore $\sup(A), \inf(A) \in f(A)$ so that
there are $x_m, x_M \in A$ such that $f(x_m) = \inf(A)$ and $f(x_M) = \sup(A)$.

\item $f:[a, b] \rightarrow \sor$ is an increasing function. Without loss of
generality, assume that $a < x_1 < \ldots, < x_n < b$. Now, 
\[
o(f, x_i) = \lim_{\delta \rightarrow 0}(M(x_i, f, \delta) - m(x_i, f, \delta))
\]
so that
\[
\sum_{i=1}^n o(f, x_i) = \lim_{\delta \rightarrow 0} \sum_{i=1}^n
(M(x_i, f, \delta) - m(x_i, f, \delta))
\]
Since $f$ is increasing,
\begin{eqnarray*}
M(x_i, f, \delta) &\le& f\left(\frac{x_i + x_{i+1}}{2}\right) \\
m(x_i, f, \delta) &\ge& f\left(\frac{x_{i-1} + x_i}{2}\right) 
\end{eqnarray*}
for any $\delta > 0$. Therefore, 
\[
\sum_{i=1}^n o(f, x_i) \le \sum_{i=1}^n f\left(\frac{x_i + x_{i+1}}{2}\right) -
f\left(\frac{x_{i-1} + x_i}{2}\right) = f\left(\frac{b + x_n}{2}\right) - 
f\left(\frac{a + x_1}{2}\right).
\]
The result follows from the fact that 
\begin{eqnarray*}
f(a) &<& f\left(\frac{a + x_1}{2}\right) \\
f(b) &>& f\left(\frac{b + x_n}{2}\right).
\end{eqnarray*}
